% !TEX root = ./report.tex

\section{Related Work}

In this section (...)\todo[inline]{To do...}

\subsection{Unread (but found interesting) papers}

\begin{itemize}
	\item ``\textbf{On building a Supercompiler for GHC}''', \\ Peter A. Jonsson
and Johan Nordlander, at Luleå University of Technology. \\
\textit{A paper describing how to build a ``Supercompiler'', which skips
intermediate representation (IR). Sounds relatively useless, but there are
quite a few papers talking about inlining in conjunction with Supercompilers.
On this list in case it becomes of relevance later on.}\\
\url{http://pure.ltu.se/portal/files/2231262/nwpt08-scp.pdf}

	\item ``\textbf{Optimizing Generics Is Easy!}'', \\ written by José Pedro
Magalhães and Stefan Holdermans and Johan Jeuring and Andres Löh, at Utrecht
University, The Netherlands. \\
\textit{Looks more like a description of how GHC inlines than anything else on
the subject of inlining.}\\
\url{http://www.andres-loeh.de/OptimizingGenerics/}

	\item ``\textbf{Should potential loop optimizations influence inlining
decisions?}'', by Christopher Barton and José Nelson Amaral and Bob Blainey, at
University of Alberta and IBM Toronto Software Laboratory, Canada. \\
\textit{Perhaps 2nd most interesting at this moment?}\\
\url{http://dl.acm.org/citation.cfm?id=961329}

	\item ``\textbf{Region-based compilation: an introduction and motivation}'',
\\ byRichard E. Hank and Wen-Mei W. Hwu, at University of Illinois, and B.
Ramakrishna Rau, at Hewlett Packard Laboratories, Palo Alto California. \\
\textit{3rd most interesting on this list at the moment.} \\
\url{http://dl.acm.org/citation.cfm?id=225160.225189&coll=DL&dl=ACM&CFID=470941312&CFTOKEN=25125832}

	\item ``\textbf{Bigloo: a portable and optimizing compiler for strict
functional languages}'', \\ by Manuel Serrano and Pierre Weis. \\
\textit{Does not really seem interesting/relevant, but Manuel Serrano has
written several relevant papers, and strict functional languages are some of
the more viable languages for this type of optimization...} \\
\url{http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.50.8424}

	\item ``\textbf{Losing functions without gaining data: another look at
defunctionalisation}'', \\ by Neil Mitchell and Colin Runciman, at University of
York, UK. \\
\textit{Not really sure what to make of this one.} \\
\url{http://dl.acm.org/citation.cfm?id=1596641}

	\item ``\textbf{Statically Unrolling Recursion to Improve Opportunities for
Parallelism}'', \\by Neil Deshpande and Stephen A. Edwards, at Columbia
University NY. \\
\textit{This one sounds cool, and being static is a plus. But that's all its got
going for it without having read it more closely. It does discuss inlining.}\\
\url{http://www.cs.columbia.edu/~sedwards/papers/deshpande2012statically.pdf}

	\item ``\textbf{An adaptive strategy for inline substitution}'', \\ by Keith
D. Cooper and Timothy J. Harvey and Todd Waterman, at Rice University Houston
and Texas Instruments Inc. Stafford, Texas. \\
\textit{Perhaps most interesting at this time due to discussing global/static
optimizations as well as being relatively new compared to other papers in this
list?}\\
\url{http://dl.acm.org/citation.cfm?id=1788381}

	\item ``\textbf{Adaptive compilation and inlining}'', \\ by Todd Waterman
and Keith D. Cooper, at Rice University Houston Texas. \\
\textit{Not sure if this is what Nico wants me to find, but this is a paper
discussing aspects of inlining I don't think I've considered much in earlier
paper searches.}\\
\url{http://dl.acm.org/citation.cfm?id=1195126}

	\item ``\textbf{Practical and effective higher-order optimizations}'', \\ by
Lars Bergstrom and Matthew Fluet and Matthew Le and John Reppy and Nora Sandler,
at Mozilla Research CA and Rochester Institute of Technology NY and University
of Chicago IL. \\
\textit{This paper discusses inlining as well as some other optimizations.
Picked for citing \cite{GHC-paper}, and being from 2014, more than much else.
I'm not sure I completely understand the abstract, but I got the feeling the
paper might have an interesting take on inlining.} \\
\url{http://dl.acm.org/citation.cfm?id=2628153}

	\item ``\textbf{Demand-driven Inlining Heuristics in a Region-based
Optimizing Compiler for ILP Architectures*}'', \\ by Tom Way and Ben Breech and
Wei Du and Lori Pollock, at University of Delaware USA. \\
\textit{This paper discusses a region-based compiling technique and how to
utilize inlining in conjunction with this, while focusing on optimizations for
ILP architectures. Picked paper for mentioning inlining in Keywords and
abstract, as well as also dealing with region based compilation which seems
relatable to Jive.} \\
\url{http://www.eecis.udel.edu/~hiper/passages/papers/waypdcsiasted01.pdf}

\end{itemize}

\subsection{Read papers}

Below papers are listed according to perceived relevance/usefulness for this
project.\\

\begin{enumerate}
	\item ``\textbf{Secrets of the Glasgow Haskell Compiler inliner}'',
\cite{GHC-paper}

This paper explains how inlining is dealt with by the GHC compiler. It spreads
this into the following sections:
\begin{enumerate}
	\item Short section describing what inlining is, and its most obvious and
generic properties and pitfalls (Section 2).
	\item An explanation of a strategy of how (and when) to inline recursive
functions (Section 3).
	\item How they so deal with name capture (Section 4).
	\item As well as discussing certain moments/situations when inlining is
carefully considered by the GHC. They also have a section describe how they
exploit their name-capture solution to support accurate tracking of both lexical
and evalutation-state environments (Section 5+6).
	\item Finally they sketch their implementation (Section 7).
\end{enumerate}

	\item ``\textbf{Inline expansion: when and how?}'',
\cite{InlineWhenHowSerrano}

This paper focuses on \textit{when} and \textit{how} to inline functions,
classified as either \textit{recursive} or \textit{non-recursive} functions.
This is done in the programming language Scheme, with \todo{confirm this?}the
Bigloo compiler M. Serrano has created.

The paper gives an algorithm for when to inline (which \cite{GHC-paper} remarks
upon), as well as comments and results on their inlining wrt. the two classes of
functions the paper defines.

	\item ``\textbf{Automatic Tuning of Inlining Heuristics}'',
\cite{AutoTuningJavaHeuristics}

This paper explores and explains the advantages of genetic algorithm  heuristics
when dynamically compiling Java. Since it is dynamically compiled, it's harder
for the compilation unit to get a proper global view of the compilation job,
required resources, and beneficial trade-offs. All of which play important roles
when deciding whether to incur the cost of increased code size when inlining.


\end{enumerate}
